---
title: "Conditional Scoring First Simulation"
output: html_notebook
---

This simulation will attempt to check whether conditionally scoring locally dependent true/false questions improves reliability in comparison to regular scoring. The steps of the simulation will include:
1. drawing 20 difficulty values for items from a uniform distribution (-3,3). assuming all to have discrimination of 1 and guessing of 0 (1PL model).
2. drawing 1000 student ability values from a standard normal distribution.
3. using the IRT ICC for each question and each student to simulate the student answering (draw number from a uniform distribution (0,1), if it's lower than the ICCi(theta) then the answer is marked as correct).
4. If the student was incorrect and it is the first of two dependent questions, then we 'flip' the score for the following item.
5. We then compare the relibility for regular scoring with two conditional scoring rules:

The strong rule - 

The weak rule - 

# Setup
```{r}
require(psych)
require(knitr)


# creating a function to simulate whether a student answered an item correctly based on the item diff and the student ability. Allowing for 1PL, 2PL, or 3PL model.
correct_response <- function(student_theta, item_b_value, item_a_value=1.7, item_c_value=0) {
  
  Pcorrect <- item_c_value + (1-item_c_value)*
    ((exp(item_a_value*(student_theta - item_b_value)))/
       (1+ exp(item_a_value*(student_theta - item_b_value))))
  
  answer <- if (runif(1) <= Pcorrect) {
  1} else {0}
  
  return(answer)
}

#Creating a function that takes a dichotomously scored table and a column number of a locally dependent item and flips the scores for that variable

flip_score <- function(Dichot, second) {
  first <- second-1
  for (i in 1:length(Dichot[1])) {
    if (Dichot[i,first]==0) {
    Dichot[i,second] <- 1-Dichot[i,second]
  } else {Dichot[i,second]}
  }
  return(Dichot)
}

condition_weak <- function(Dichot, second) {
  first <- second-1
  for (i in 1:length(Dichot[1])) {
    if (Dichot[i,first]==0 & Dichot[i,second]==0) {
      Dichot[i,second] <- 1
    }else {Dichot[i,second]}
    
  }
  return(Dichot)
}

condition_strong <- function(Dichot, second) {
  first <- second-1
  for (i in 1:length(Dichot[1])) {
    if (Dichot[i,first]==0) {
      Dichot[i,second] <- 1-Dichot[i,second]
    }else {Dichot[i,second]}
    
  }
  return(Dichot)
}

```

#Drawing the variables
```{r}
item_diff <- runif(20,-3,3)
theta <- rnorm(1000)
```


# Creating the dichotomous response table
```{r}
Dichot <- data.frame(matrix(NA, nrow=length(theta), ncol=length(item_diff)))

# dichot table when all items are independent
for (i in 1:length(theta)) {
  for (j in 1:length(item_diff)) {
    Dichot[i,j] <- correct_response(theta[i],item_diff[j])
  }
}

Alpha <- psych::alpha(Dichot)
Alpha <- as.numeric(Alpha$total[1])
```


# Inserting locally dependent items

```{r}
dependent_items <- c(2,6,10,11,15,20) # Item 1 cannot be dependent

Dichot_dep <- Dichot

# flipping scores for all locally dependent items
for (i in dependent_items) {
  Dichot_dep <- flip_score(Dichot_dep,i)
  
}

#Alpha with dependencies
Alpha_dep <- psych::alpha(Dichot_dep)
Alpha_dep <- as.numeric(Alpha_dep$total[1])
```


# Conditional scoring - weak condition
```{r}
Dichot_cond_weak <- Dichot_dep

for (i in dependent_items) {
  Dichot_cond_weak <- condition_weak(Dichot_cond_weak,i)
  
}

Alpha_cond_weak <- psych::alpha(Dichot_cond_weak)
Alpha_cond_weak <- as.numeric(Alpha_cond_weak$total[1])


```


#Conditional scoring - strong condition
```{r}
Dichot_cond_strong <- Dichot_dep

for (i in dependent_items) {
  Dichot_cond_strong <- condition_strong(Dichot_cond_strong,i)
  
}

Alpha_cond_strong <- psych::alpha(Dichot_cond_strong)
Alpha_cond_strong <- as.numeric(Alpha_cond_strong$total[1])

```


# Display simulation results

```{r}
alpha_table <- as.tibble(t(c(Alpha,Alpha_dep,Alpha_cond_weak,Alpha_cond_strong)))
names(alpha_table) <- c("independent","dependent","weak condition","strong condition")
knitr::kable(alpha_table,"rst")

```

The strong condition seems to "fix" the local dependence (since it was created with a strong assumption) and therefore always has the same alpha value and the independent sample.
Dependent test forms seem to always have lower reliability than independent test forms.

The weak condition, however, is interesting since it sometimes improves reliability but at worse, has the same reliability as the weak condition.

To really see these trends, I will run the simulation multiple times and look at the average alpha values for all runs.

# Now do it s times for differnt "occasions" (same items, same students), it can take a long while, 500 runs took me about 10 minutes

```{r}
#Creating fixed values and empty df
sims <- 100 #number of simulations
item_diff <- runif(20,-3,3)
theta <- rnorm(1000)
dependent_items <- c(2,6,10,11,15,20) # arbitrarily chosen, item 1 cannot be dependent
Alphas_occasions <- data.frame(matrix(NA, nrow=sims, ncol=4)) #collecting the alpha results from each round 

for (s in 1:sims) {
  
  Dichot <- data.frame(matrix(NA, nrow=length(theta), ncol=length(item_diff)))

# dichot table when all items are independent
for (i in 1:length(theta)) {
  for (j in 1:length(item_diff)) {
    Dichot[i,j] <- correct_response(theta[i],item_diff[j])
  }
}

Alpha <- psych::alpha(Dichot)
Alpha <- as.numeric(Alpha$total[1])

Dichot_dep <- Dichot

# flipping scores for all locally dependent items
for (i in dependent_items) {
  Dichot_dep <- flip_score(Dichot_dep,i)
  
}

#Alpha with dependencies
Alpha_dep <- psych::alpha(Dichot_dep)
Alpha_dep <- as.numeric(Alpha_dep$total[1])

Dichot_cond_weak <- Dichot_dep

for (i in dependent_items) {
  Dichot_cond_weak <- condition_weak(Dichot_cond_weak,i)
  
}

Alpha_cond_weak <- psych::alpha(Dichot_cond_weak)
Alpha_cond_weak <- as.numeric(Alpha_cond_weak$total[1])
Dichot_cond_strong <- Dichot_dep

for (i in dependent_items) {
  Dichot_cond_strong <- condition_strong(Dichot_cond_strong,i)
  
}

Alpha_cond_strong <- psych::alpha(Dichot_cond_strong)
Alpha_cond_strong <- as.numeric(Alpha_cond_strong$total[1])

Alphas_occasions[s,] <- t(c(Alpha,Alpha_dep,Alpha_cond_weak,Alpha_cond_strong))
}

names(Alphas_occasions) <- c("independent","dependent","weak condition","strong condition")

colMeans(Alphas_occasions)
```
First run of 500 simulations results in:

independent        dependent      weak condition  strong condition 
  0.7562680        0.7561136        0.7562170        0.7561727 

it seems that the general trend is 

r(dependent) <= r(strong_condition) <= r(weak_condition) <= r(independent)


