---
title: "Conditional Scoring - Simulation 2 - correlated abilities and test saavy"
output: html_notebook
---

In this I will attempt to create a computationaly efficient way to draw N theta values (standard normal), I item difficulties (standard normal) and discriminations (Gamma(2,2)). 
See https://ben18785.shinyapps.io/distribution-zoo/ to explore Gamma (2,2) PDF and CDF.
In the first phase I will assume that one item pair is locally dependent.

I will compare two cases: 
1. the first item measures one ability (theta1) and item two measures a second ability (theta2) normally distributed and highly correlated (r=.75 or so).
2. both items measure the same ability

Trying two flipping mechanisms:
1. naive - fully deterministic.
2. test-saavy based - draw a value [0,1] negatively correlated with ability, to simulate the test-savvy of students. If students are test-savvy, they are unlikely to follow the previous question in answering the current. Students incorrect on the first question will be tested for on their test-savvy, if they fail, we flip their answer to the second question.

#Setup and basic variables
```{r, echo=FALSE, warning=FALSE}
library(tibble)
library(psych)
library(knitr)
library(irtoys)
library(boot)
library(base)
```


#Drawing functions
## Students abilities
Creating a table of student ability with a correlated second ability (for one variation of the model) and a test-savvy characteristic correlated with the average of the two ability values (for another variation).
```{r}


inv.logit <- sigmoid <- function(z) {
	1/(1+exp(-z))
}

logistic2pl.vec <- function(sk, d, a) {
        ### uses vectors and returns matrix
	smat = matrix(rep(sk,length(d)),nrow=length(sk))
	dmat = matrix(rep(d,length(sk)),nrow=length(sk),byrow=T)
	amat = matrix(rep(a,length(sk)),nrow=length(sk),byrow=T)	
	sigmoid(amat * (smat - dmat))
}


draw_student_ability <- function(students = 10, rho_ability=0.75) {
  # Student ability - Math ability
  theta1 <- rnorm(students) 
  # Avoiding misconcept, correlated with first
  theta2 <- theta1*rho_ability+sqrt(1-rho_ability^2)*rnorm(students)
  #creating tibble
  theta <- cbind(theta1,theta2)
  theta <- as_tibble(theta)
  return(theta)
}
```

## Item Characteristics
Drawing item characteristics. difficulty is drawn from a standard normal distribution. Discrimination is drawn from a Gamma distribution with a shape of 2 and rate of 2. this means discrimination values are postive with a mean of 1, and the distribution is right-skewed which seemed more reasonable to me than other options.
```{r}
draw_item_characteristics <- function(items = 20) {
  #item difficulty
  difficulty <- rnorm(items)
  #item discrimination - range [0,inf], mean 1.
  discrimination <- rgamma(items,shape = 2, rate = 2)
  # creating tibble
  items <- cbind(difficulty,discrimination)
  items <- as_tibble(items)
  return(items)
}
```

## Dichotomous response table

```{r}
## THIS FUNCTION HAD A LOT OF ISSUES
# simulate data from an IRT model, uni-dimensional
Draw_dichot_table <- function(ability,item_diff,item_disc) {
  nitems <- length(item_diff)
  nstudents <- length(ability)
  # a*theta-b/a  OR a*(theta-b) ## THIS LOOKS WRONG
  probability_matrix <- inv.logit(outer(ability,item_disc,FUN = "*")-(item_diff*item_disc))
  # creating table of same dimensions with random values drawn from unif[0,1]
 ## THIS IS WRONG random <- matrix(rnorm(length(ability)*length(item_diff)), ncol=length(item_diff))
  
  # creating the dichotomous response table
  dichot_table <- (runif(nitems*nstudents) < probability_matrix) + 0
  return(dichot_table)
}

## I think misconcept chance should also related with diff and disc, but misconcept chance should be less than incorrect, so add 20% to the probability matrix. 
Draw_misconcept_table <- function(misconcept,item_diff,item_disc) {
  nitems <- length(item_diff)
  nstudents <- length(ability)
  # a*theta-b/a  OR a*(theta-b) ## THIS LOOKS WRONG
  probability_matrix <- inv.logit(outer(misconcept,item_disc,FUN = "*")-(item_diff*item_disc))
  # creating table of same dimensions with random values drawn from unif[0,1]
 ## THIS IS WRONG random <- matrix(rnorm(length(ability)*length(item_diff)), ncol=length(item_diff))
  
  # creating the dichotomous response table
  misconcept_table <- (runif(nitems*nstudents) < (probability_matrix+0.15)) + 0
  return(misconcept_table)
}

Decrease_misconcept_table <- function(ability,misconcept_result,item_diff,item_disc) {
  nitems <- length(item_diff)
  nstudents <- length(ability)
  # a*theta-b/a  OR a*(theta-b) ## THIS LOOKS WRONG
  probability_matrix <- inv.logit(outer(ability,item_disc,FUN = "*")-(item_diff*item_disc))
  # creating table of same dimensions with random values drawn from unif[0,1]
 ## THIS IS WRONG random <- matrix(rnorm(length(ability)*length(item_diff)), ncol=length(item_diff))
  
  # creating the dichotomous response table
  correct_table <- (runif(nitems*nstudents) < (probability_matrix+0.2*(misconcept_result-1))) + 0
  misconcept_table <- (runif(nitems*nstudents) < probability_matrix) + 0
  return(misconcept_table + correct_table)
}

Absolute_misconcept_table <- function(ability,misconcept_result,item_diff,item_disc) {
  nitems <- length(item_diff)
  nstudents <- length(ability)
  # a*theta-b/a  OR a*(theta-b) ## THIS LOOKS WRONG
  probability_matrix <- inv.logit(outer(ability,item_disc,FUN = "*")-(item_diff*item_disc))
  # creating table of same dimensions with random values drawn from unif[0,1]
 ## THIS IS WRONG random <- matrix(rnorm(length(ability)*length(item_diff)), ncol=length(item_diff))
  
  # creating the dichotomous response table
  correct_table <- (runif(nitems*nstudents) < (probability_matrix+misconcept_result-1)) + 0
  misconcept_table <- (runif(nitems*nstudents) < probability_matrix) + 0
  return(misconcept_table + correct_table)
}
```

## Get results for use

```{r}
students = 20
ability <- draw_student_ability(students = 20,0.75)
item1 <- draw_item_characteristics(1)
#we can also try random misconcept
#here 40% students would have misconcept
random_misconcept <- (runif(students) < 0.4) + 0
#misconcept happens to q1
question1_misconcept <- Draw_misconcept_table(ability$theta2,item1$difficulty,item1$discrimination)
#ability of q1
question1_ability <- Draw_dichot_table(ability$theta1,item1$difficulty,item1$discrimination)
#decrease certain amount probability if misconcept happens
question1_decrease <- Decrease_misconcept_table(ability$theta1,question1_misconcept,item1$difficulty,item1$discrimination)
#if misconcept happens, answer is wrong
question1_absolute <- Absolute_misconcept_table(ability$theta1,question1_misconcept,item1$difficulty,item1$discrimination)

item2 <- draw_item_characteristics(1)
question2_misconcept <- Draw_misconcept_table(ability$theta2,item2$difficulty,item2$discrimination)
question2_ability <- Draw_dichot_table(ability$theta1,item2$difficulty,item2$discrimination)
question2_decrease <- Decrease_misconcept_table(ability$theta1,question1_misconcept,item2$difficulty,item2$discrimination)
#if misconcept happens, answer is wrong
question2_absolute <- Absolute_misconcept_table(ability$theta1,question2_misconcept,item2$difficulty,item2$discrimination)
```

## First Model

# Misconcept happens on both questions or neither. If misconcept happens, probability of correct reduced by certain amount.
# 0-wrong answer, 1-wrong because of misconcept, 2-correct

```{r}
model1 <- cbind(question1_decrease,question2_decrease)
```

## Second Model

# Misconcept happens on both questions or neither. If misconcept happens, the answer would be wrong.


```{r}
model2 <- cbind(question1_absolute,question2_absolute)
```


## Third Model

# If misconcept happens on question, question is wrong. Otherwise, look at ability.
# If misconcept happens on question1, 90% chance happens to question2. If not, 10% chance for question2.

```{r}
change_happen <- (runif(students) < 0.1) + 0
misconcept <- cbind(question1_misconcept,change_happen)
misconcept[misconcept[,2]==1,1] <- 1-misconcept[misconcept[,2]==1,1]
question2_absolute <- Absolute_misconcept_table(ability$theta1,misconcept[,1],item2$difficulty,item2$discrimination)
model3 <- cbind(question1_absolute,question2_absolute)
```

